### 2018-08-14 sections/signal_background_modeling.tex

Signal and most background processes are modelled using Monte Carlo (MC) simulations.
After the event preselection, the main background is $\ttbar$ production, often in association with jets, denoted by $\ttbar$+jets in the following.
Small contributions arise from single-top-quark, $W/Z$+jets, multijet and diboson ($WW,WZ,ZZ$) production, as well as from the associated 
production of a vector boson $V$ ($V=W,Z$) or a Higgs boson and a $\ttbar$ pair ($\ttbar V$ and $\ttbar H$). All backgrounds 
with prompt leptons, i.e. those originating from the decay of a $W$ boson, a $Z$ boson, or a $\tau$-lepton,
are estimated using samples of simulated events and initially normalised to their theoretical cross sections.
In the simulation, the top quark and SM Higgs boson masses are set to $172.5~\gev$ and $125~\gev$, respectively,
and the Higgs boson is allowed to decay into all SM particles with branching ratios calculated using \textsc{Hdecay}~\cite{Djouadi:1997yw}.  
Backgrounds with non-prompt electrons or muons, with photons or jets misidentified as electrons, or with jets misidentified as $\tauhad$ candidates, 
generically referred to as ``fake leptons", are estimated using data-driven methods. 
The background prediction is further improved during the statistical analysis by performing a likelihood 
fit to data using several signal-depleted search regions, as discussed in Sections~\ref{sec:strategy_Hbb} and~\ref{sec:strategy_Htautau}.

@@

Samples of simulated $\ttbar \to WbHq$ events are generated with the next-to-leading-order (NLO) generator\footnote{In the following, 
the order of a generator should be understood as referring to the order in the strong coupling constant at which the matrix element calculation 
is performed.} {\amcatnlolong}~2.4.3~\cite{Alwall:2014hca}  (referred to in the following as {\amcatnlo}) with the NNPDF3.0~\cite{Ball:2014uwa} parton distribution function (PDF) set and interfaced to {\pythia} 8.212~\cite{Sjostrand:2007gs} for the modelling of parton showering, hadronisation, and the underlying event. 
The A14~\cite{ATLASUETune4} set of tuned parameters in {\pythia} controlling the description of multiparton interactions and  
initial and final state radiation, referred to as the ``tune'', is used.
The signal sample is normalised to the same cross section as used for the $t\bar{t}\to WbWb$ sample (see discussion below) and
assuming a reference branching ratio $\BR_{\mathrm{ref}}(t\to Hq)=1\%$.
The case of both top quarks decaying into $Hq$ is neglected in the analysis given existing upper limits on $\BR(t \to Hq)$ (see Section~\ref{sec:intro}).

The nominal sample used to model the $\ttbar$ background is generated with the NLO generator {\powheg}~v2 \cite{Frixione:2007nw,Nason:2004rx,Frixione:2007vw,Alioli:2010xd} using the NNPDF3.0 PDF set. The {\powheg} model parameter $h_{\textrm{damp}}$, which controls 
matrix element to parton shower matching and effectively regulates the high-$\pt$ radiation, is set to 1.5 times the top quark mass. 
Parton shower and hadronisation are modelled by {\pythia}~8.2 with the A14 tune.
Alternative $\ttbar$ simulation samples used to derive systematic uncertainties are described in Section~\ref{sec:syst_bkgmodeling}. 
The generated $\ttbar$ samples are normalised to a theoretical cross section of $832^{+46}_{-51}$~pb, 
computed using \textsc{Top++}~v2.0~\cite{Czakon:2011xx} at next-to-next-to-leading order (NNLO), 
including resummation of next-to-next-to-leading logarithmic (NNLL) soft gluon 
terms~\cite{Cacciari:2011hy,Baernreuther:2012ws,Czakon:2012zr,Czakon:2012pz,Czakon:2013goa}, 
and using the MSTW 2008 NNLO~\cite{Martin:2009iq,Martin:2009bu} PDF set. 
The theoretical uncertainties result from variations of the factorisation and renormalisation scales, as well as from uncertainties in the
top quark mass, and in the PDF and $\alpha_{\textrm{S}}$. The latter two represent the largest contribution to the overall theoretical uncertainty 
in the cross section and were calculated using the PDF4LHC prescription~\cite{Botje:2011sn} 
with the MSTW 2008 68\% CL NNLO, CT10 NNLO~\cite{Lai:2010vv,Gao:2013xoa} and NNPDF2.3 5f FFN~\cite{Ball:2012cx} PDF sets.

The $\Hbb$ search requires a more sophisticated treatment of the $\ttbar$ background. Therefore, the simulated $\ttbar$ events are categorised depending 
on the flavour content of additional particle jets not originating from the decay of the $\ttbar$ system (see Ref.~\cite{Aaboud:2017rss} for details).
Events labelled as either \ttbin\ or \ttcin\ are generically referred to in the following as $\ttbar$+HF events, where HF stands for ``heavy flavour''.
The remaining events are labelled as $\ttbar$+light-jets events, including those with no additional jets. 
A finer categorisation of \ttbin\ events is considered for the purpose of applying further corrections and
assigning systematic uncertainties associated with the modelling of heavy-flavour production in different topologies~\cite{Aaboud:2017rss}.
In particular, the \ttbin\ events are reweighted to an NLO prediction in the four-flavour (4F) scheme 
of \ttbin\ production including parton showering~\cite{Cascioli:2013era}, based on {\ShOLlong}~\cite{Gleisberg:2008ta, Cascioli:2011va} (referred to as {\ShOL} 
in the following) using the CT10 PDF set.  This reweighting is performed in such a way that the inter-normalisations of the \ttbin\ categories are at NLO accuracy, 
while preserving the nominal \ttbin\ cross section in {\powheg}+{\pythia}. This reweighting is applied as well to the alternative $\ttbar$ samples.

Samples of single-top-quark events corresponding to the $t$-channel production mechanism are generated with the 
{\powheg}~v1~\cite{Frederix:2012dh} generator that uses the 4F scheme  for the NLO matrix element calculations
and the fixed 4F \textsc{CT10}f\textsc{4}~\cite{Lai:2010vv} PDF set.
Samples corresponding to the $Wt$- and $s$-channel production mechanisms are generated 
with {\powheg}~v1 using the CT10 PDF set. Overlaps between the $\ttbar$ and $Wt$ final states are avoided by using 
the ``diagram removal'' scheme~\cite{Frixione:2005vw}.
The parton shower, hadronisation and the underlying event are modelled using {\pythia}~6.428 with the CTEQ6L1 PDF set 
in combination with the Perugia 2012 tune~\cite{Skands:2010ak}.
The single-top-quark samples are normalised to the approximate NNLO theoretical cross 
sections~\cite{Kidonakis:2011wy,Kidonakis:2010ux,Kidonakis:2010tc}.

@@

Samples of $WW/WZ/ZZ$+jets events are generated with {\sherpa}~2.1.1 using the CT10 PDF set
and include processes containing up to four electroweak vertices. The matrix element includes zero additional partons 
at NLO and up to three partons at LO using the same procedure as for the $W/Z$+jets samples. 
The final states simulated require one of the bosons to decay leptonically and the other hadronically.
All diboson samples are normalised to their NLO theoretical cross sections provided by {\sherpa}.

Samples of $\ttbar V$ and $\ttbar H$ events are generated with {\amcatnlo}~2.2.1, using 
NLO matrix elements and the NNPDF3.0 NLO~\cite{Ball:2014uwa} PDF set. 
Showering is performed using {\pythia}~8.210 and the A14 tune.
The $\ttbar V$ samples are normalised to the NLO cross section computed with {\amcatnlo}.
The $\ttbar H$ sample is normalised using the NLO cross section~\cite{Raitio:1978pt,Beenakker:2002nc,Dawson:2003zu,Yu:2014cka,Frixione:2015zaa}  
and the Higgs boson decay branching ratios calculated using \textsc{Hdecay}.

@@

The fake $\tauhad$ background prediction in a given SR is modelled by the fake template derived from its corresponding CR, which is a data-driven method. The fake template is defined as the data subtracted by the contributions from the simulated backgrounds with real $\tauhad$ candidates. In the $\lephad$ channel, simulation studies indicate that the fake $\tauhad$ background composition is consistent between the SR and the CR, and dominated by $\ttbar$. In this case, the fake template is simply normalised to the fake $\tauhad$ background yield in the SR and taken as the fake $\tauhad$ background prediction. In the $\hadhad$ channel, the fake $\tauhad$ background is expected to be dominated by multijet production. However, simulation studies indicate that the fraction of $\ttbar$ yield is higher in the SR than in the CR. Therefore, an appropriate amount of simulated $\ttbar$ events with fake $\tauhad$ candidates in the CR are added to the fake template to match the fake $\tauhad$ background composition in the SR. This modified fake template is then normalised to the expected fake $\tauhad$ background yield in the SR, as sone in the case of the $\lephad$ channel.

>>>>

Signal and most background processes are modelled using Monte Carlo (MC) simulation.
After the event preselection, the main background is $\ttbar$ production, often in association with jets, denoted by $\ttbar$+jets in the following.
Small contributions arise from single-top-quark, $W/Z$+jets, multijet and diboson ($WW,WZ,ZZ$) production, as well as from the associated 
production of a vector boson $V$ ($V=W,Z$) or a Higgs boson and a $\ttbar$ pair ($\ttbar V$ and $\ttbar H$). All backgrounds 
with prompt leptons, i.e. those originating from the decay of a $W$ boson, a $Z$ boson, or a $\tau$-lepton,
are estimated using samples of simulated events and initially normalised to their theoretical cross sections.
In the simulation, the top quark and SM Higgs boson masses are set to $172.5~\gev$ and $125~\gev$, respectively,
and the Higgs boson is allowed to decay into all SM particles with branching ratios calculated using \textsc{Hdecay}~\cite{Djouadi:1997yw}.  
Backgrounds with non-prompt electrons or muons, with photons or jets misidentified as electrons, or with jets misidentified as $\tauhad$ candidates, 
generically referred to as ``fake leptons", are estimated using data-driven methods. 
The background prediction is further improved during the statistical analysis by performing a likelihood 
fit to data using several signal-depleted analysis regions, as discussed in Sections~\ref{sec:strategy_Hbb} and~\ref{sec:strategy_Htautau}.

@@

Samples of simulated $\ttbar \to WbHq$ events are generated with the next-to-leading-order (NLO) generator\footnote{In the following, 
the order of a generator should be understood as referring to the order in the strong coupling constant at which the matrix element calculation 
is performed.} {\amcatnlolong}~2.4.3~\cite{Alwall:2014hca}  (referred to in the following as {\amcatnlo}) with the NNPDF3.0 NLO~\cite{Ball:2014uwa} parton distribution function (PDF) set and interfaced to {\pythia} 8.212~\cite{Sjostrand:2007gs} with the NNPDF2.3 LO~\cite{Ball:2012cx} PDF set for the modelling of parton showering, hadronisation, and the underlying event. 
The A14~\cite{ATLASUETune4} set of tuned parameters in {\pythia} controlling the description of multiparton interactions and  
initial and final state radiation, referred to as the ``tune'', is used.
The signal sample is normalised to the same cross section as used for the $t\bar{t}\to WbWb$ sample (see discussion below) and
assuming a reference branching ratio $\BR_{\mathrm{ref}}(t\to Hq)=1\%$.
The case of both top quarks decaying into $Hq$ is neglected in the analysis given existing upper limits on $\BR(t \to Hq)$ (see Section~\ref{sec:intro}).

The nominal sample used to model the $\ttbar$ background is generated with the NLO generator {\powheg}~v2 \cite{Frixione:2007nw,Nason:2004rx,Frixione:2007vw,Alioli:2010xd} using the NNPDF3.0 NLO PDF set. The {\powheg} model parameter $h_{\textrm{damp}}$, which controls 
matrix element to parton shower matching and effectively regulates the high-$\pt$ radiation, is set to 1.5 times the top quark mass. 
Parton shower, hadronisation, and underlying event are modelled by {\pythia}~8.210 with the NNPDF2.3 LO PDF set in combination with the A14 tune.
Alternative $\ttbar$ simulation samples used to derive systematic uncertainties are described in Section~\ref{sec:syst_bkgmodeling}. 
The generated $\ttbar$ samples are normalised to a theoretical cross section of $832^{+46}_{-51}$~pb, 
computed using \textsc{Top++}~v2.0~\cite{Czakon:2011xx} at next-to-next-to-leading order (NNLO), 
including resummation of next-to-next-to-leading logarithmic (NNLL) soft gluon 
terms~\cite{Cacciari:2011hy,Baernreuther:2012ws,Czakon:2012zr,Czakon:2012pz,Czakon:2013goa}, 
and using the MSTW 2008 NNLO~\cite{Martin:2009iq,Martin:2009bu} PDF set. 
The theoretical uncertainties result from variations of the factorisation and renormalisation scales, as well as from uncertainties in the
top quark mass, and in the PDF and $\alpha_{\textrm{S}}$. The latter two represent the largest contribution to the overall theoretical uncertainty 
in the cross section and were calculated using the PDF4LHC prescription~\cite{Botje:2011sn} 
with the MSTW 2008 68\% CL NNLO, CT10 NNLO~\cite{Lai:2010vv,Gao:2013xoa} and NNPDF2.3 5f FFN~\cite{Ball:2012cx} PDF sets.

%The $\Hbb$ search requires a more sophisticated treatment of the $\ttbar$ background. 
The $\ttbar$ background selected by the $\Hbb$ search is enriched in $\ttbar$+heavy-flavour production, and thus requires a more sophisticated 
treatment than provided by the nominal $\ttbar$ sample. 
%The large and complex $\ttbar$ background selected by the $\Hbb$ search requires a more sophisticated treatment than provided by the nominal $\ttbar$ sample. 
Therefore, the simulated $\ttbar$ events are categorised depending 
on the flavour content of additional particle jets not originating from the decay of the $\ttbar$ system (see Ref.~\cite{Aaboud:2017rss} for details).
Events labelled as either \ttbin\ or \ttcin\ are generically referred to in the following as $\ttbar$+HF events, where HF stands for ``heavy flavour''.
The remaining events are labelled as $\ttbar$+light-jets events, including those with no additional jets. 
A finer categorisation of \ttbin\ events is considered for the purpose of applying further corrections and
assigning systematic uncertainties associated with the modelling of heavy-flavour production in different event topologies~\cite{Aaboud:2017rss}.
In particular, the \ttbin\ events are reweighted to an NLO prediction in the four-flavour (4F) scheme 
of \ttbin\ production including parton showering~\cite{Cascioli:2013era}, based on {\ShOLlong}~\cite{Gleisberg:2008ta, Cascioli:2011va} (referred to as {\ShOL} 
in the following) using the CT10 PDF set.  This reweighting is performed in such a way that the inter-normalisations of the \ttbin\ categories are at NLO accuracy, 
while preserving the nominal \ttbin\ cross section of the nominal $\ttbar$ sample. This reweighting is applied to the alternative $\ttbar$ samples as well.

Samples of single-top-quark events corresponding to the $t$-channel production mechanism are generated with the 
{\powheg}~v1~\cite{Frederix:2012dh} generator that uses the 4F scheme  for the NLO matrix element calculations
and the fixed 4F \textsc{CT10}f\textsc{4}~\cite{Lai:2010vv} PDF set.
Samples corresponding to the $tW$- and $s$-channel production mechanisms are generated 
with {\powheg}~v1 using the CT10 PDF set. Overlaps between the $\ttbar$ and $tW$ final states are avoided by using 
the ``diagram removal'' scheme~\cite{Frixione:2005vw}.
The parton shower, hadronisation and the underlying event are modelled using {\pythia}~6.428 with the CTEQ6L1~\cite{Pumplin:2002vw,Nadolsky:2008zw} PDF set 
in combination with the Perugia 2012 tune~\cite{Skands:2010ak}.
The single-top-quark samples are normalised to the approximate NNLO theoretical cross 
sections~\cite{Kidonakis:2011wy,Kidonakis:2010ux,Kidonakis:2010tc}.

@@

Samples of $WW/WZ/ZZ$+jets events are generated with {\sherpa}~2.1.1 using the CT10 PDF set
and include processes containing up to four electroweak vertices. 
In the case of $WW/WZ$+jets ($ZZ$+jets) the matrix element is calculated for zero (up to one) additional partons 
at NLO and up to three partons at LO using the same procedure as for the $W/Z$+jets samples. 
The final states simulated require one of the bosons to decay leptonically and the other hadronically.
All diboson samples are normalised to their NLO theoretical cross sections provided by {\sherpa}.

Samples of $\ttbar V$ and $\ttbar H$ events are generated with {\amcatnlo}~2.2.1, using NLO matrix elements and the NNPDF3.0 NLO PDF set,
and interfaced to {\pythia}~8.210 with the NNPDF2.3 LO PDF set and the A14 tune.
The $\ttbar V$ samples are normalised to the NLO cross section computed with {\amcatnlo}.
The $\ttbar H$ sample is normalised using the NLO cross section~\cite{Raitio:1978pt,Beenakker:2002nc,Dawson:2003zu,Yu:2014cka,Frixione:2015zaa}.

@@

The fake $\tauhad$ background prediction in a given SR is modelled by the distribution (referred to as ``fake $\tauhad$ template") derived from data in the corresponding CR. The fake $\tauhad$ template is defined as the data distribution from which the contributions from the simulated backgrounds with real $\tauhad$ candidates are subtracted. In the $\lephad$ channel, simulation studies indicate that the fake $\tauhad$ background composition is consistent between the SR and the CR, and dominated by $\ttbar$. In the $\hadhad$ channel, the fake $\tauhad$ background is expected to be dominated by multijet production. However, simulation studies indicate that the contribution of $\ttbar$ events to the fake $\tauhad$ background is higher in the SR than in the CR. Therefore, an appropriate amount of simulated $\ttbar$ events with fake $\tauhad$ candidates in the CR is added to the fake $\tauhad$ template to match the fake $\tauhad$ background composition in the SR. 
In both the $\lephad$ and $\hadhad$ channels, the fake $\tauhad$ template in each SR is initially normalised to the estimated fake $\tauhad$ background yield, 
defined as the data yield minus the contributions from the simulated backgrounds with real $\tauhad$ candidates (i.e. assuming no signal contribution).
During the statistical analysis, the normalisation of the fake $\tauhad$ background in each SR will be allowed to vary unconstrained in the fit to data, as discussed in Section~\ref{sec:results_Htautau}.
